---
title: "Petro Analysis"
author: "Shelby Ziegler"
date: "2023-09-12"
output: html_document
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=FALSE, fig.width=14, fig.height=10, out.width = "75%")

rm(list = ls());graphics.off()
library(tidyverse) # For data manipulation and summarization
library(plotrix) #To use standard error calculation
library(cowplot) #For grouping plots
library(scales) #For editing axis scales
library(vegan) #For multivariate analyses
library(rstatix) #For running ANCOVA
library(ggpmisc) #adding equation and r2 to figures
library(viridis)
library(ggpubr)
library(visreg)
library(MuMIn)


alpha=0.05
formula<-y~x
theme_set(theme_classic(base_size=16))
```


Examine the environmental correlates with the invasive green porcelain crab. *Including 9 sites within Georgia Coastal Ecosystem LTER domain.

``````{r} 
petro<-read.csv("Data/Petro/AllPetroDataCombined.csv")

str(petro)

petro$Cluster<-factor(petro$Cluster, levels=c("Tybee", "Romerly Marsh", "Little Ogeechee", "Liberty-Sunbury", "McIntosh", "High Point", "Meridian", "Dean Creek", "McKay River", "Glynn-Jointer Creek", "Camden"))

p1<-petro%>%
  dplyr::select(Latitude, Rugosity, Total.oyster.density, Total.mudcrab.density, Petro.density, Temp, Salinity, Marsh.water.ratio, Nearest.Inlet, Nearest.road, Neartest.Armored.Shoreline)

corrplot::corrplot(cor(p1),method="number", type="upper")

#model diagnostics
library(vcd)
ggplot(petro, aes(Petro.density)) + geom_histogram() 

#Use Ord_plot to determine distribution type for models
pfit<-goodfit(petro$Petro.density, type="nbinomial", method="MinChisq" )
plot(pfit, main="nbinomial")
summary(pfit) #Weird output (due to NAs produced - verify with JMP)

Ord_plot(petro$Petro.density) #Tells us negative binomial is the best distribution.

```

Visually examine relationships between petro and other variables

```{r}
library(MASS)

#Latitude
ggplot(petro, aes(x=Latitude, y=Petro.density))+
  geom_point(shape=21,size=3, aes(fill=Cluster))+
  scale_fill_viridis_d()+
  geom_smooth(method="glm.nb", color="black")+
  labs(x="Latitude", y=expression(paste(italic(Petrolithes~armatus)~density~(0.25~m^2))))+
  theme(panel.border = element_rect(color = "black", fill=NA))

#Rugosity
ggplot(petro, aes(x=Rugosity, y=Petro.density))+
  geom_point(shape=21,size=3, aes(fill=Cluster))+
  scale_fill_viridis_d()+
  geom_smooth(method="glm.nb", color="black")+
  labs(x="Rugosity", y=expression(paste(italic(Petrolithes~armatus)~density~(0.25~m^2))))+
  theme(panel.border = element_rect(color = "black", fill=NA))

#Oyster density
ggplot(petro, aes(x=Total.oyster.density, y=Petro.density))+
  geom_point(shape=21,size=3, aes(fill=Cluster))+
  scale_fill_viridis_d()+
  geom_smooth(method="glm.nb", color="black")+
  labs(x="Oyster density", y=expression(paste(italic(Petrolithes~armatus)~density~(0.25~m^2))))+
  theme(panel.border = element_rect(color = "black", fill=NA))

#mud crab density
ggplot(petro, aes(x=Total.mudcrab.density, y=Petro.density))+
  geom_point(shape=21,size=3, aes(fill=Cluster))+
  scale_fill_viridis_d()+
   geom_smooth(method="glm.nb", color="black")+
  labs(x="Mud crab density", y=expression(paste(italic(Petrolithes~armatus)~density~(0.25~m^2))))+
  theme(panel.border = element_rect(color = "black", fill=NA))

#Salinity
ggplot(petro, aes(x=Salinity, y=Petro.density))+
  geom_point(shape=21,size=3, aes(fill=Cluster))+
  scale_fill_viridis_d()+
  geom_smooth(method="glm.nb", color="black")+
  labs(x="Salinity", y=expression(paste(italic(Petrolithes~armatus)~density~(0.25~m^2))))+
  theme(panel.border = element_rect(color = "black", fill=NA))

#Temp
ggplot(petro, aes(x=Temp, y=Petro.density))+
  geom_point(shape=21,size=3, aes(fill=Cluster))+
  scale_fill_viridis_d()+
  geom_smooth(method="glm.nb", color="black")+
  labs(x="Temperature", y=expression(paste(italic(Petrolithes~armatus)~density~(0.25~m^2))))+
  theme(panel.border = element_rect(color = "black", fill=NA))

#Inlet
ggplot(petro, aes(x=Nearest.Inlet, y=Petro.density))+
  geom_point(shape=21,size=3, aes(fill=Cluster))+
  scale_fill_viridis_d()+
  geom_smooth(method="glm.nb", color="black")+
  labs(x="Nearest Ocean Inlet (km)", y=expression(paste(italic(Petrolithes~armatus)~density~(0.25~m^2))))+
  theme(panel.border = element_rect(color = "black", fill=NA))

#Road
ggplot(petro, aes(x=Nearest.road, y=Petro.density))+
  geom_point(shape=21,size=3, aes(fill=Cluster))+
  scale_fill_viridis_d()+
  geom_smooth(method="glm.nb", color="black")+
  labs(x="Nearest road (km)", y=expression(paste(italic(Petrolithes~armatus)~density~(0.25~m^2))))+
  theme(panel.border = element_rect(color = "black", fill=NA))

#Marsh water ratio
ggplot(petro, aes(x=Marsh.water.ratio, y=Petro.density))+
  geom_point(shape=21,size=3, aes(fill=Cluster))+
  scale_fill_viridis_d()+
  geom_smooth(method="glm.nb", color="black")+
  labs(x="Marsh:Water", y=expression(paste(italic(Petrolithes~armatus)~density~(0.25~m^2))))+
  theme(panel.border = element_rect(color = "black", fill=NA))

#Freshwater input
ggplot(petro, aes(x=Freshwater.input, y=Petro.density))+
  geom_point(shape=21,size=3, aes(fill=Cluster))+
  scale_fill_viridis_d()+
  geom_smooth(method="glm.nb", color="black")+
  labs(x="Nearest freshwater input", y=expression(paste(italic(Petrolithes~armatus)~density~(0.25~m^2))))+
  theme(panel.border = element_rect(color = "black", fill=NA))

#Altamaha
ggplot(petro, aes(x=Linear.dist.from.Altamaha, y=Petro.density))+
  geom_point(shape=21,size=3, aes(fill=Cluster))+
  scale_fill_viridis_d()+
  geom_smooth(method="glm.nb", color="black")+
  labs(x="Distance from largest freshwater input (Altamaha)", y=expression(paste(italic(Petrolithes~armatus)~density~(0.25~m^2))))+
  theme(panel.border = element_rect(color = "black", fill=NA))
```

Conduct negative binomial models 

```{r}
#Models 
library(lme4)
library(MASS)


#First examine relationship with latitude
petrolat<-MASS::glm.nb(Petro.density~Latitude,  data=petro, na.action=na.fail)
summary(petrolat) ## Petro positively correlated
r.squaredGLMM(petrolat)
#Examine if the latitude model is a better predictor of Petro than random chance (null model) using the analysis of deviance.
anova(petrolat)
car::Anova(petrolat) #checking second method for deviance anlaysis

#Visualize the model output for petro and latitude
visreg(petrolat)


#Run full model for petro.
petroall<-MASS::glm.nb(Petro.density~scale(Salinity)+scale(Rugosity)+scale(Total.oyster.density)+scale(Total.mudcrab.density)+scale(Nearest.Inlet)+scale(Nearest.road),  data=petro, na.action=na.fail)

ggplot(petro, aes(x= Total.mudcrab.density, y=Petro.density))+
  geom_point()

#Use the dredge function from MuMIn to examine all possible model combinations.
x1<-dredge(petroall, "partial.sd", extra =list(R2 = function(x)     {
  r.squaredGLMM(x, null = nullmodel)["delta", ]}))
#Subset out models with delta AIC < 2
subset(x1,delta <= 2, recalc.weights = FALSE)

#Conduct model averaging to determine the most important variables in the top models.
summary(model.avg(subset(x1,delta <= 2, recalc.weights = FALSE)))


## R ^2 range 0.23-0.29x
###model averaging --- mud crabs top fit. 

#Examine mud crab only model and look at visualization of model output with visreg
p1<-MASS::glm.nb(Petro.density~Total.mudcrab.density,  data=petro, na.action=na.fail)
 visreg(p1, ylab="P(Petro density)",type="conditional", scale="response", partial=TRUE)
 
#Examine mud crab and rugosity model and look at visualization of model output with visreg 
p2<-MASS::glm.nb(Petro.density~scale(Total.mudcrab.density)+scale(Rugosity),  data=petro, na.action=na.fail)

summary(p2)
par(mfrow=c(1,2))
 visreg(p2, ylab="P(Petro density)",type="conditional", scale="response", partial=TRUE )
 par(mfrow=c(1,1))
``` 
 
 
 
```{r} 
#Remove potential outliers and run models again to ensure robustness. Remove mud crab density > 140 per 0.25m2^2
petrosub<-petro%>%
  filter(Total.mudcrab.density<140)

petrosuball<-MASS::glm.nb(Petro.density~scale(Salinity)+scale(Rugosity)+scale(Total.oyster.density)+scale(Total.mudcrab.density)+scale(Nearest.Inlet)+scale(Nearest.road),  data=petrosub, na.action=na.fail)

x2<-dredge(petrosuball, "partial.sd", extra =list(R2 = function(x)     {
  r.squaredGLMM(x, null = nullmodel)["delta", ]}))
subset(x2,delta <= 2, recalc.weights = FALSE)

summary(model.avg(subset(x2,delta <= 2, recalc.weights = FALSE)))

### Mud crab still come out as top model (R2 = 0.18)

#Run best fit model for visualization
ps1<-MASS::glm.nb(Petro.density~Total.mudcrab.density,  data=petrosub, na.action=na.fail)
 
#Visualize output
visreg(ps1, ylab="P(Petro density)",type="conditional", scale="response", partial=TRUE)

```

Try running models with large and small mud crabs separated out

```{r}
# Data fram with large (>10mm) and small (<10mm) mud crabs in separate columns for 24 of 33 reefs 
x<-read.csv("Data/CIGOysterAllCombinedWideFormat.csv")%>%
  dplyr::select(Site, large.mudcrab.density, Small.mudcrab.density)


## Crabs from the three sapelo clusters did not get separated into large and small. Using measurements from each sample we calcualted the proportion of the total sample that would be large vs small and then created large vs small categories. 
mud2<-petro%>%
  filter(Cluster=="High Point" | Cluster=="Meridian" |Cluster=="Dean Creek")
mp<-read.csv("Data/MP-mudprop.csv")

mud3<-left_join(mud2, mp, by="Site")

mud3$large.mudcrab.density<-(mud3$Total.mudcrab.density*(1-mud3$Prop))
mud3$Small.mudcrab.density<-(mud3$Total.mudcrab.density*(mud3$Prop))

mud3$large.mudcrab.density<-round(mud3$large.mudcrab.density)
mud3$Small.mudcrab.density<-round(mud3$Small.mudcrab.density)

mud3<-mud3%>%
  dplyr::select(Site, large.mudcrab.density, Small.mudcrab.density)

#Combine calculated large vs small data with the rest of the sites. 
mudx<-rbind(x, mud3)

petro<-left_join(petro, mudx, by="Site")


#rerun model with large and small mudcrabs 
petroall1<-MASS::glm.nb(Petro.density~scale(Salinity)+scale(Rugosity)+scale(Total.oyster.density)+scale(large.mudcrab.density)+scale(Small.mudcrab.density)+scale(Nearest.Inlet)+scale(Nearest.road),  data=petro, na.action=na.fail)

x3<-dredge(petroall1, "partial.sd", extra =list(R2 = function(x)     {
  r.squaredGLMM(x, null = nullmodel)["delta", ]}))

subset(x3,delta <= 2, recalc.weights = FALSE)

summary(model.avg(subset(x3,delta <= 2, recalc.weights = FALSE)))

sm1<-MASS::glm.nb(Petro.density~scale(Small.mudcrab.density),  data=petro, na.action=na.fail)
lg1<-MASS::glm.nb(Petro.density~scale(large.mudcrab.density),  data=petro, na.action=na.fail)

visreg(sm1)
visreg(lg1)
```


Calculate mean Petro lengths

```{r}
sp<-read.csv("Data/ReefSpecies-2023-02-06.csv")

psp<-sp%>%
  filter(species.ds=="petro")%>%
  select(Site, c(X1:X20))%>%
  gather(Count, Size, -Site)

pspm<-psp%>%
  summarize(meanp=mean(Size, na.rm=T), se=std.error(Size, na.rm=T))
```