---
title: "Petro Analysis"
author: "Shelby Ziegler"
date: "2023-09-12"
output: html_document
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=FALSE, fig.width=14, fig.height=10, out.width = "75%")

rm(list = ls());graphics.off()
library(tidyverse) # For data manipulation and summarization
library(plotrix) #To use standard error calculation
library(cowplot) #For grouping plots
library(scales) #For editing axis scales
library(vegan) #For multivariate analyses
library(rstatix) #For running ANCOVA
library(ggpmisc) #adding equation and r2 to figures
library(viridis)
library(ggpubr)
library(visreg)
library(MuMIn)


alpha=0.05
formula<-y~x
theme_set(theme_classic(base_size=16))
```


Examine the environmental correlates with the invasive green porcelain crab. *Including 3 sites within GCE domain


``````{r}
petro<-read.csv("Data/Petro/AllPetroDataCombined.csv")

str(petro)

petro$Cluster<-factor(petro$Cluster, levels=c("Tybee", "Romerly Marsh", "Little Ogeechee", "Liberty-Sunbury", "McIntosh", "High Point", "Meridian", "Dean Creek", "McKay River", "Glynn-Jointer Creek", "Camden"))

p1<-petro%>%
  dplyr::select(Latitude, Rugosity, Total.oyster.density, Total.mudcrab.density, Petro.density, Temp, Salinity, Marsh.water.ratio, Nearest.Inlet, Nearest.road, Neartest.Armored.Shoreline)

corrplot::corrplot(cor(p1),method="number", type="upper")
```

Visually examine relationships between petro and other variables

```{r}
library(MASS)

#Latitude
ggplot(petrosub, aes(x=Latitude, y=Petro.density))+
  geom_point(shape=21,size=3, aes(fill=Cluster))+
  scale_fill_viridis_d()+
  geom_smooth(method="glm.nb", color="black")+
  labs(x="Latitude", y=expression(paste(italic(Petrolithes~armatus)~density~(0.25~m^2))))+
  theme(panel.border = element_rect(color = "black", fill=NA))

#Rugosity
ggplot(petro, aes(x=Rugosity, y=Petro.density))+
  geom_point(shape=21,size=3, aes(fill=Cluster))+
  scale_fill_viridis_d()+
  geom_smooth(method="glm.nb", color="black")+
  labs(x="Rugosity", y=expression(paste(italic(Petrolithes~armatus)~density~(0.25~m^2))))+
  theme(panel.border = element_rect(color = "black", fill=NA))

#Oyster density
ggplot(petro, aes(x=Total.oyster.density, y=Petro.density))+
  geom_point(shape=21,size=3, aes(fill=Cluster))+
  scale_fill_viridis_d()+
  geom_smooth(method="glm.nb", color="black")+
  labs(x="Oyster density", y=expression(paste(italic(Petrolithes~armatus)~density~(0.25~m^2))))+
  theme(panel.border = element_rect(color = "black", fill=NA))

#mud crab density
ggplot(petro, aes(x=Total.mudcrab.density, y=Petro.density))+
  geom_point(shape=21,size=3, aes(fill=Cluster))+
  scale_fill_viridis_d()+
  geom_smooth(method="glm.nb", color="black")+
  labs(x="Mud crab density", y=expression(paste(italic(Petrolithes~armatus)~density~(0.25~m^2))))+
  theme(panel.border = element_rect(color = "black", fill=NA))

#Salinity
ggplot(petro, aes(x=Salinity, y=Petro.density))+
  geom_point(shape=21,size=3, aes(fill=Cluster))+
  scale_fill_viridis_d()+
  geom_smooth(method="glm.nb", color="black")+
  labs(x="Salinity", y=expression(paste(italic(Petrolithes~armatus)~density~(0.25~m^2))))+
  theme(panel.border = element_rect(color = "black", fill=NA))

#Temp
ggplot(petro, aes(x=Temp, y=Petro.density))+
  geom_point(shape=21,size=3, aes(fill=Cluster))+
  scale_fill_viridis_d()+
  geom_smooth(method="glm.nb", color="black")+
  labs(x="Temperature", y=expression(paste(italic(Petrolithes~armatus)~density~(0.25~m^2))))+
  theme(panel.border = element_rect(color = "black", fill=NA))

#Inlet
ggplot(petro, aes(x=Nearest.Inlet, y=Petro.density))+
  geom_point(shape=21,size=3, aes(fill=Cluster))+
  scale_fill_viridis_d()+
  geom_smooth(method="glm.nb", color="black")+
  labs(x="Nearest Ocean Inlet (km)", y=expression(paste(italic(Petrolithes~armatus)~density~(0.25~m^2))))+
  theme(panel.border = element_rect(color = "black", fill=NA))

#Road
ggplot(petro, aes(x=Nearest.road, y=Petro.density))+
  geom_point(shape=21,size=3, aes(fill=Cluster))+
  scale_fill_viridis_d()+
  geom_smooth(method="glm.nb", color="black")+
  labs(x="Nearest road (km)", y=expression(paste(italic(Petrolithes~armatus)~density~(0.25~m^2))))+
  theme(panel.border = element_rect(color = "black", fill=NA))

#Marsh water ratio
ggplot(petro, aes(x=Marsh.water.ratio, y=Petro.density))+
  geom_point(shape=21,size=3, aes(fill=Cluster))+
  scale_fill_viridis_d()+
  geom_smooth(method="glm.nb", color="black")+
  labs(x="Marsh:Water", y=expression(paste(italic(Petrolithes~armatus)~density~(0.25~m^2))))+
  theme(panel.border = element_rect(color = "black", fill=NA))

#Freshwater input
ggplot(petro, aes(x=Freshwater.input, y=Petro.density))+
  geom_point(shape=21,size=3, aes(fill=Cluster))+
  scale_fill_viridis_d()+
  geom_smooth(method="glm.nb", color="black")+
  labs(x="Nearest freshwater input", y=expression(paste(italic(Petrolithes~armatus)~density~(0.25~m^2))))+
  theme(panel.border = element_rect(color = "black", fill=NA))

#Altamaha
ggplot(petro, aes(x=Linear.dist.from.Altamaha, y=Petro.density))+
  geom_point(shape=21,size=3, aes(fill=Cluster))+
  scale_fill_viridis_d()+
  geom_smooth(method="glm.nb", color="black")+
  labs(x="Distance from largest freshwater input (Altamaha)", y=expression(paste(italic(Petrolithes~armatus)~density~(0.25~m^2))))+
  theme(panel.border = element_rect(color = "black", fill=NA))
```


```{r}
#Models 
library(lme4)
library(MASS)

petroX<-petro%>%
  filter(Petro.density>3)%>%
  summarize(meanp=mean(Petro.density), sep=std.error(Petro.density))

petrosub<-petro%>%
  filter(Petro.density>3 & Total.mudcrab.density<140)


#Look at relationship with latitude
petrolat<-MASS::glm.nb(Petro.density~Latitude,  data=petro, na.action=na.fail)
summary(petrolat)
anova(petrolat)
car::Anova(petrolat)
visreg(petrolat)

#Bio petro model
petrobio<-MASS::glm.nb(Petro.density~scale(Salinity)+scale(Rugosity)+scale(Total.oyster.density)+scale(Total.mudcrab.density),  data=petro, na.action=na.fail)
summary(petrobio)

summary(MASS::glm.nb(Petro.density~Rugosity, data=petrosub))

dredge(petrobio,"sd", extra =list(R2 = function(x)     {
  r.squaredGLMM(x, null = nullmodel)["delta", ]}))


#Top 2 models - Mud crab (r2=0.23), mud crab + rugosity (r=0.26)


#Landscape petro model
petroland<-MASS::glm.nb(Petro.density~scale(Marsh.water.ratio)+scale(Nearest.Inlet)+scale(Nearest.road)+scale(Freshwater.input),  data=petro, na.action=na.fail)
summary(petroland)


dredge(petroland,"sd", extra =list(R2 = function(x)     {
  r.squaredGLMM(x, null = nullmodel)["delta", ]}))


#Combine landscape and biological data 
petroall<-MASS::glm.nb(Petro.density~scale(Salinity)+scale(Rugosity)+scale(Total.oyster.density)+scale(Total.mudcrab.density)+scale(Nearest.Inlet)+scale(Nearest.road),  data=petro, na.action=na.fail)

x1<-dredge(petroall, "partial.sd", extra =list(R2 = function(x)     {
  r.squaredGLMM(x, null = nullmodel)["delta", ]}))
subset(x1,delta <= 2, recalc.weights = FALSE)

summary(model.avg(subset(x1,delta <= 2, recalc.weights = FALSE)))


###model averaging --- mud crabs. 

p1<-MASS::glm.nb(Petro.density~Total.mudcrab.density,  data=petrosub, na.action=na.fail)
 visreg(p1, ylab="P(Petro density)",type="conditional", scale="response", partial=TRUE)
 
p2<-MASS::glm.nb(Petro.density~scale(Total.mudcrab.density)+scale(Rugosity),  data=petro, na.action=na.fail)

summary(p2)
par(mfrow=c(1,2))
 visreg(p2, ylab="P(Petro density)",type="conditional", scale="response", partial=TRUE )
 par(mfrow=c(1,1))
 
```

